Constraints on the mass and occupation fraction of dark matter halos hosting Lyman-alpha emitters at $z\sim 3$

### 2013-06-26 paper/ms.tex
little changes to the paper

Some theoretical approaches to this problem have been based on
ab-initio approach modeling. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models  and  full N-body hidrodynamical simulations
\citep{Laursen2007, Dayal2009, ForeroRomero2011, Yajima2012, Orsi2012}.

Added to the uncertainties in the astrophysical processes describing
star formation in galactic populations, a highly debated step in this
approach is the calculation of the fraction of \ly photons
that escape the galaxy to the observer. Given the resonant nature of
the \lya line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM) \citep{Neufeld1991, ForeroRomero2011, Dijkstra2012,
Laursen2013}.

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that effectively host a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ y $f_{\rm occ}$.

@@

The second benchmark is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA instead of $190$\AA
used by \citep{Yamada2012}

@@

\begin{equation}
\xi(\theta) = \left(\frac{r}{\theta_{0}}\right)^{-\gamma}
\label{eq:fitting}
\end{equation}

@@

This result provides the basis to understand why only a specific range of models
${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models
with a minimum mass $\log_{10} M_{\rm min}>11.5$\hMsun will always have a
surface number density lower than the observational
constrain. This makes impossible that models with that minimum mass
can be compatible with observations.

@@

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. The results are very similar between the {\texttt
  match} and {\texttt random} methods. With this constraints the
number of consistent models with  $\log_{10}M_{\rm min}< 11.7$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that there is not a
strong selection effect on the occupation fraction $f_{\rm occ}$.

@@

In order to better illustrate the results from this test we
show in the right panel of Figure \ref{figure:correlation_parameters} 
the plane $\theta_{0}$-$M_{\rm min}$ for all models. There we divide
the models into two disjoint sets: those with $\log \frac{M_{\rm
    max}}{M_{\odot}}<12$ (blue dots) and $\log \frac{M_{\rm
    max}}{M_{\odot}}>12$ (green dots). The colored rectangle includes
the parameter region which is consistent with the observational
constraint in $\theta_{0}$. With this restriction we find that most
models with $\log \frac{M_{\rm max}}{M_{\odot}}>12$ and  $\log
\frac{M_{\rm     min}}{M_{\odot}}>11.1$ can be safely ruled out.

@@

We find that to make a physical interpretation of these models it is
useful to do it in terms of the size of halo mass range. We now use the
variable $\Delta M=\log_{10}M_{\rm max} - \log M_{\rm  min}$, together
with the escape fraction $f_{\rm occ}$ to help us build a classification
of all the successful models into three families:

@@

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the total absence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun shows that our models support theoretical insights
where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}.

@@

Using abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} report that the instantaneous star
formation efficiency (star formation rate divided by the stellar mass)
presents a clear maximum around $10^{11.7}$\Msun at all redshifts
$z<4$.

This mass scale is strictly superior to the great majority of  $M_{\rm max}$
values allowed in the models with low escape fraction. Using the results published
in \cite{Behroozi2013a} we find that the typical stellar mass in halos
of $10^{11.4}$\hMsun is $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while
their star formation rate is in the range $0.6\pm 0.2$ Msun yr$^{-1}$.
This halo mass range around $10^{11.4}$ \hMsun is the lower bound of
what is computed in the abundance matching model in order to fit the
observational data for the stellar mass function based on the
observations of Lyman Break Galaxies.

@@

All the models also present a range of masses where the halos are
below the mass scale of $10^{11.4}$\hMsun inferred as a lower
threshold for the LBGs used in halo abundance matching
studies. Detailed study of high-z LAES will have a positive impact on
our understanding of the connection of low mass halos to star forming
galaxies.

>>>>

Some theoretical approaches to this problem have been based on
ab-initio approach modeling. Starting from the DM halo population, the
corresponding intrinsic star formation properties are inferred
together with other statistics such as the luminosity function, the
correlation function and the equivalent width distributions. Such
modelling has been implemented from analytic considerations,
semi-analytic models  and  full N-body hidrodynamical simulations
\citep{Laursen2007, Dayal2009, ForeroRomero2011, Yajima2012, Orsi2012, Soler2012}.

Added to the uncertainties in the astrophysical processes describing
star formation in galactic populations, a highly debated step in this
approach is the calculation of the fraction of \ly photons
that escape the galaxy to the observer. Given the resonant nature of
the \lya line, the radiative transfer of \ly is sensitive to the
density, temperature, topology and kinematics of the neutral Hydrogen
in the interstellar medium (ISM) \citep{Neufeld1991, ForeroRomero2011, Dijkstra2012,
Laursen2013,Orsi2012}.

@@

In our model a dark matter halo can only host one or zero LAE. 
There are three parameters in the model that decide whether a halo can host a
galaxy or not: the lower and upper bounds for the mass range $M_{\rm min}<
M_{\rm h} < M_{\rm max}$ where LAEs reside and the fraction $f_{\rm
  occ}$ of such halos that effectively host a LAE. We do not assign a
luminosity to each LAE. We are primarily interested in
constraining the halo mass range hosting detectable LAEs under the
conditions defined by \cite{Yamada2012}. In what follows we will describe by
the letter ${\mathcal M}$ a model defined by an specific choice of the
three scalar parameters $M_{\rm min}$, $M_{\rm  max}$ y $f_{\rm occ}$.

@@

The second benchmark is the angular correlation function (ACF). The
measurements presented in \citep{Yamada2012} do not report an ACF
measurement. Instead we use the results reported by
\citep{Hayashino2004} on the LAEs observed in the densest field of
SSA22. It is important to keep in mind that there are some differences
between this work and \citep{Yamada2012}. First, the color selection by
\cite{Yamada2012} is less stringent compared to the one by
\cite{Hayashino2004}. Also the EW threshold is different,
\cite{Hayashino2004} uses a cut around $154$\AA \ instead of $190$\AA
\ used by \citep{Yamada2012}

@@

\begin{equation}
\xi(\theta) = \left(\frac{\theta}{\theta_{0}}\right)^{-\gamma}
\label{eq:fitting}
\end{equation}

@@

This result provides the basis to understand why only a specific range of models
${\mathcal M}$ can be expected to be consistent with
observations. From Figure \ref{fig:halos} we can read that models
with a minimum mass $\log_{10} M_{\rm min}>11.5$\hMsun \ will always have a
surface number density lower than the observational
constrain. This makes impossible that models with that minimum mass
can be compatible with observations.

@@

Figure \ref{fig:restriction_mock} presents the locii of these models
in the parameter space $M_{\rm min}-M_{\rm max}$ and $M_{\rm
  min}-f_{\rm occ}$. The results are very similar between the {\texttt
  match} and {\texttt random} methods. With this constraints the
number of consistent models with  $\log_{10}M_{\rm min}< 11.5$ are
greatly reduced. This corresponds to the regions in the parameter
space in Figure \ref{fig:landscape} that already had a low number of
consistent mock surveys. On the other hand, from the right panel in
Figure \ref{fig:restriction_mock} one can see that there is not a
strong selection effect on the occupation fraction $f_{\rm occ}$.

@@

In order to better illustrate the results from this test we
show in the right panel of Figure \ref{figure:correlation_parameters} 
the plane $\theta_{0}$-$M_{\rm min}$ for all models. There we divide
the models into two disjoint sets: those with $\log_{10} M_{\rm
    max}<12$ (blue dots) and $\log_{10} M_{\rm
    max}>12$ (green dots). The colored rectangle includes
the parameter region which is consistent with the observational
constraint in $\theta_{0}$. With this restriction we find that most
models with $\log_{10} M_{\rm max}>12$ and  $\log_{10} M_{\rm min}>11.1$ 
%can be safely ruled out.

@@

We find that to make a physical interpretation of these models it is
useful to do it in terms of the size of halo mass range. We now use the
variable $\Delta M=\log_{10}M_{\rm max} - \log_{10}M_{\rm  min}$, together
with the escape fraction $f_{\rm occ}$ to help us build a classification
of all the successful models into three families:

@@

The preference for narrow $\Delta M$ ranges to hosting LAEs together
with the total absence of reasonable models with $M_{\rm max}>
10^{12}$\hMsun \  shows that our models support theoretical insights
where the most massive systems are not bright \ly sources
\citep{ForeroRomero2012}.

@@

Using abundance matching methods throughout cosmic time from redshifts
$0<z<8$ \cite{Behroozi2013a,Behroozi2013b} report that the instantaneous star
formation efficiency (star formation rate divided by the stellar mass)
presents a clear maximum around $10^{11.7}$\Msun \ at all redshifts
$z<4$.

This mass scale is strictly superior to the great majority of  $M_{\rm max}$
values allowed in the models with low escape fraction. Using the results published
in \cite{Behroozi2013a} we find that the typical stellar mass in halos
of $10^{11.4}$\hMsun \ is $(1.0\pm0.3)\times 10^{9.0}$ \hMsun, while
their star formation rate is in the range $0.6\pm 0.2$ Msun yr$^{-1}$.
This halo mass range around $10^{11.4}$ \hMsun \ is the lower bound of
what is computed in the abundance matching model in order to fit the
observational data for the stellar mass function based on the
observations of Lyman Break Galaxies.

@@

All the models also present a range of masses where the halos are
below the mass scale of $10^{11.4}$\hMsun \  inferred as a lower
threshold for the LBGs used in halo abundance matching
studies. Detailed study of high-z LAES will have a positive impact on
our understanding of the connection of low mass halos to star forming
galaxies.
